Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-08-02 Paper/assa.tex
added a result on fixing unimportant params ...

It was shown in~\cite{Lamboni:2013} that the total Sobol' 
index $T_i(f)$ can be bounded in terms of $\nu_i(f)$:
\begin{equation}\label{equ:sobol_bound}
T_i(f) \leq \frac{C_i}{\V(f)}\nu_i(f), \quad i = 1, \ldots, \Np,
\end{equation}
where for each $i$, $C_i$ is an appropriate \emph{Poincar\'{e}} constant.
This provides a strong theoretical basis for using DGSMs to identify 
unimportant inputs. Moreover, 
in~\cite{Vohra:2018} we introduced the normalized screening metric
\[
   \tilde{\nu}_i(f) = \frac{C_i \nu_i(f)}{\sum_{i=1}^\Np C_i \nu_i(f)},
\]
for detecting unimportant input parameters. 
Note that if $\xi_i$ are iid, the $C_i$'s will cancel and the normalized screening
metric reduces to 
\[
    \tilde{\nu}_i(f) = \frac{\nu_i(f)}{\sum_{i=1}^\Np \nu_i(f)} 
      = \frac{\vec{e}_i^T \mat{C} \vec{e}_i}{\trace(\mat{C})} 
      = \frac{\sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2}{\sum_{k = 1}^\Np \lambda_k}.
\]
\newcommand{\I}{\mathcal{I}}
Combining Proposition~\ref{prp:dgsm_bound} and~\eqref{equ:sobol_bound}, shows
an interesting link between the activity scores and total Sobol' indices.
Specifically, by computing the activity scores, we can indetify unimportant
parameters.

One can attempt to reduce parameter dimension by fixing
unimportant parameters at nominal values.

%Let $\I \subset \{1, \ldots,
%Np\}$ be an index set that indexes the unimportant parameters in the parameter
%vector $\vec{\xi}$.  Letting $|\I|$ denote the number of elements of $\I$, 
%we denote by $\vec{z} \in \R^{|\I|}$ a generic vector of nominal values for
%the unimportant parameters $\vec\xi_\I$, and let $\vec{y}$ be a generic vector of
%important variables, $\vec\xi_{\I^c}$,
%where $\I^c$ denotes the complement of $\I$ in $\{1, \ldots, \Np\}$. 
Consider the reduced model
The \emph{reduced} model, obtained by fixing unimportant variables is given by
\newcommand{\xin}{\vec{\xi}^{(\sim i)}}
\[
f_{z}(\vec\xi) = f(\xin, z)
\] 
where $(\vec{y}, \vec{z})$ should be understood as a vector $\vec\xi$ with
$\vec\xi_{I^c}$ set to $\vec{y}$ and $\vec\xi_{\I}$ set to $\vec{z}$.
The following result quantifies the approximation error when using $f_\vec{z}$
in place of $f$:
\begin{theorem}
Consider the relative error
\[
\delta(z) =
\frac{ \int_\Omega \big( f(\vec\xi) - f_{z}(\vec\xi)\big)^2 \, \mu(d\vec\xi) }
          {\int_\Omega \big( f(\vec\xi) )^2 \, \mu(d\vec\xi)}.
\] 
We have
\[
          \Ez{ \delta(z)} \leq \frac{2 C_i}{\V(f)}\big(\act{f}{i}{r} + \lambda_{r+1}\big).
\]
\end{theorem}
%
%\[
%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
%\]
%we obtain the following useful inequality, which
%links activity scores and Sobol' indices:
%for $1 \leq r \leq \Np$, and $i = 1, \ldots, \Np$, we have
%\[
%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
%\]
%This was noted in~\cite{Diaz:2016} for the specific case of all $\xi_i$'s being
%iid $U(0, 1)$, in which case $C_i = 1/\pi^2$, $i = 1, \ldots, \Np$.

@@

>>>>

It was shown in~\cite{Lamboni:2013} that the total Sobol' 
index $T_i(f)$ can be bounded in terms of $\nu_i(f)$:
\begin{equation}\label{equ:sobol_bound}
T_i(f) \leq \frac{C_i}{\V(f)}\nu_i(f), \quad i = 1, \ldots, \Np,
\end{equation}
where for each $i$, $C_i$ is an appropriate \emph{Poincar\'{e}} constant.
This provides a strong theoretical basis for using DGSMs to identify 
unimportant inputs. Moreover, 
in~\cite{Vohra:2018} we introduced the normalized screening metric
\[
   \tilde{\nu}_i(f) = \frac{C_i \nu_i(f)}{\sum_{i=1}^\Np C_i \nu_i(f)},
\]
for detecting unimportant input parameters. 
Note that if $\xi_i$ are iid, the $C_i$'s will cancel and the normalized screening
metric reduces to 
\[
    \tilde{\nu}_i(f) = \frac{\nu_i(f)}{\sum_{i=1}^\Np \nu_i(f)} 
      = \frac{\vec{e}_i^T \mat{C} \vec{e}_i}{\trace(\mat{C})} 
      = \frac{\sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2}{\sum_{k = 1}^\Np \lambda_k}.
\]

@@

Combining Proposition~\ref{prp:dgsm_bound} and~\eqref{equ:sobol_bound}, shows
an interesting link between the activity scores and total Sobol' indices.
Specifically, by computing the activity scores, we can indetify unimportant
parameters.  
One can attempt to reduce parameter dimension by fixing
unimportant parameters at nominal values. Suppose $\xi_i$ is
deemed unimportant as a result of a global sensitivity analysis
based on computing the activity scores. We want to understand
what is the approximation error that occurs once we fix $\xi_i$ 
at a nominal value.
%Let $\I \subset \{1, \ldots,
%Np\}$ be an index set that indexes the unimportant parameters in the parameter
%vector $\vec{\xi}$.  Letting $|\I|$ denote the number of elements of $\I$, 
%we denote by $\vec{z} \in \R^{|\I|}$ a generic vector of nominal values for
%the unimportant parameters $\vec\xi_\I$, and let $\vec{y}$ be a generic vector of
%important variables, $\vec\xi_{\I^c}$,
%where $\I^c$ denotes the complement of $\I$ in $\{1, \ldots, \Np\}$. 
%For a vector $\xi$, we let $\xin$ denote the vector obtained by removing
%$\xi_i$.

Let $\vec\xi$ be given and let $z$ be a nominal value for $\xi_i$;  we define 
$\vec{y}^z(\vec\xi)$ as the vector with entries $y^z_j = \xi_j$ for $j \neq i$
and $y^z_i = z_i$.
We consider the \emph{reduced} model, 
obtained by fixing $\xi_i$ at a nominal value $z \in \R$: 
\[
f^{z}(\vec\xi) = f(\vec{y}^z(\vec\xi))
\] 
We consider the following error indicator.
\[
\delta(z) =
\frac{ \int_\Omega \big( f(\vec\xi) - f^{z}(\vec\xi)\big)^2 \, \mu(d\vec\xi) }
          {\int_\Omega f(\vec\xi)^2 \, \mu(d\vec\xi)}.
\]

\begin{theorem}
We have $\Ez{ \delta(z)} \leq 2C_i\big(\act{f}{i}{r} + \lambda_{r+1}\big)/{\V(f)} $.
\end{theorem}
%
%\[
%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
%\]
%we obtain the following useful inequality, which
%links activity scores and Sobol' indices:
%for $1 \leq r \leq \Np$, and $i = 1, \ldots, \Np$, we have
%\[
%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
%\]
%This was noted in~\cite{Diaz:2016} for the specific case of all $\xi_i$'s being
%iid $U(0, 1)$, in which case $C_i = 1/\pi^2$, $i = 1, \ldots, \Np$.
